### 2021-09-30 papers/modeling-of-combined-lfr-csp-sco2-paper/content.tex
removed comments

\subsection{Cycle Component Modeling}
%======================================================================================
Components present in the cycles are modeled using various techniques and are discussed in more detail below. Turbines and compressors are analyzed using isentropic efficiencies. Counter-flow heat exchangers are modelled using the effectiveness-NTU method while simplified ``black box'' heat exchangers that use a simple energy balance for state point calculations are used in lieu of more detailed component models where data is available. The lead-cooled fast reactor is assumed to be a black box heat exchanger because of the constant heat input and state points on the sCO$_2$ inlet and outlet are provided. The molten salt loop for the CSP is modeled with necessary components including hot and cold TES, receiver, pumps and counter-flow heat exchangers.

@@

\mw{this is kind of an awkward way to describe what's happening. I think most readers are familiar with isentropic efficiencies and this can be omitted.-BW}

@@

\mw{This is for a single flow stream. Why would there be $\dot{Q}_{in}$ and $\dot{Q}_{out}$ in a single flow stream? Also, the variable should be consistent with the figure, which would be $\dot{Q}_{HX}$. -BW}

@@

\mw{If mass is conserved, use a single $\dot{m}$ - BW}

@@

\mw{How do we know that 400C is the optimal temperature? This information is given to us by Westinghouse, right? -BW}

@@

\mw{Best practice is to allow figures to float. To do this, specify the location as [htb] ("here, top, or bottom"). I've done a search-replace for all figures to do this. -BW}

@@

The C2S heat exchanger is active in the 'charging' cycle operating modes, when the focus is on heat storage for later use. Pump 1 is actively moving solar salt from cold TES to hot TES through the C2S heat exchanger extracting heat from the sCO$_2$ Brayton cycle. Additionally, while the focus is on heat storage, and the heliostat field is inputting heat, pump 2 is actively transporting solar salt through the receiver to be stored in the hot TES.
\mw{Is this necessarily true? Couldn't there be a condition where we'd like to dump energy from the LFR while simultaneously collecting from CSP?-BW}

@@

\subsection{Standardization of Cycle Modeling}
%======================================================================================
\mw{What do you mean by ``generalized?''-BW}

@@

In addition to standardized parameters, all cycles have identical recompression sides. The recompression side contains a precooler, low temperature recuperator, and two compressors; main compressor and recompressor. 
\mw{We need an overview here of our process for identifying which cycles will be modeled, as the discussion below gets a bit complicated with all of the different permutations. It might be helpful to have a model taxonomy that schematically shows how the different models are related to each other.-BW}
\mw{components don't need to be capitalized -- they aren't proper nouns -BW}
The modeled cycles are summarized in Table \ref{tab-cycle_sum}.

@@

The cycle that is using the LFR heat input in the two-cycle configuration is labeled as C-LFR-ON and the cycle diagram is illustrated in Figure \ref{c-lfr-on}. 
\mw{There are lots of little 1-2 sentence ``paragraphs'' that could be consolidated with some thought.-BW}

@@

\bl{be clear that Salt HX links back into the sCO2 cycle (e.g. parallel to 5-6) but is not shown to simplify the diagram. Otherwise people will wonder where it is coming from-BW}
\mw{After lots of tinkering, I've not solved how to get the figures to avoid spilling over the edge of the page. I think it's a problem with the MDPI template. The figures have a lot of whitespace between components and numbers, and you may need to try to compress some of them down to make them more compact.-BW}
\end{paracol}

@@

The C2S is located around the turbine and LFR at state points 5 to 6 depending on the charging cycle configuration: pre, parallel, post or circulator. In the C-1HTR1T-ON cycle, the LFR HX and CSP HX have identical inlet temperatures due to splitting the flow prior to their parallel orientation. Therefore, three sensitivity studies are done on the model. The initial two studies have the low LFR temperature constrained to the value of $400^{\circ}$C with varied cold CSP TES temperature and maximized cycle efficiency. To achieve a maximum cycle efficiency, the split fraction amount of flow to the main compressor, $y_{1}$, is parametrically studied.
\mw{State what variable(s) are manipulated in order to maximize cycle efficiency -BW} 
Two cold TES temperatures are tested with constrained LFR low temperature of $400^{\circ}$C: 
\begin{itemize}
    \item	$410^{\circ}$C: Lowest cold TES temperature possible due to the sCO$_2$ cold inlet constrained from the LFR to $400^{\circ}$C and the addition of $10^{\circ}$C approach temperature;
    \item	$440^{\circ}$C: Upper bound temperature on cold TES storage;
    \item	$390^{\circ}$C: Unconstrained LFR cold inlet temperature allows a lower cold TES of $390^{\circ}$C to be achieved. This allows for a larger temperature drop across the CSP HX, increasing dispatchability.
\end{itemize}

@@

This cycle, C-2HTR3T-ON, can be seen in Figure \ref{c-2htr3t-on} and has two high temperature recuperators and three turbines. The LFR is powering one turbine, T1, and transferring unused heat to the flow entering LFR HX through a dedicated high temperature recuperator, HTR. The cycle with heat addition from the CSP powers the other two turbines, which for modelling purposes are combined into a single turbine T2, while having a dedicated high temperature recuperator, HTR2. The two turbines displayed as T2 can be modeled as a singular turbine because their isentropic efficiencies are identical causing the inlet and outlet conditions of the turbines to be consistent with a singular turbine. Additionally, the power produced by the two turbines is proportional to mass flow rate, each receives a fraction of the mass flow rate therefore producing the same fraction of power. Summing these power fractions together yields the total power of a singular turbine in the same position. 
\mw{Need to explain how T2 is actually 2 separate turbines represented diagramatically as 1, and why we can model it this way.-BW} 
After the high temperature recuperators, the two flows are combined and sent to the LTR hot side.

@@

Charging cycle configurations accommodate energy storage modes of operation. These configurations examine the location of LFR heat extraction via C2S. To maximize the available heat for extraction, alternator net power is set to zero, therefore requiring the turbine power to balance with the compressors' demand. Despite the components being non-ideal and consuming power, the recompression cycle continues to operate, ensuring that there is mass flow to transfer heat from the Brayton cycle to C2S. The excess energy from the LFR is thermally stored in the TES for later use when grid demand increases. Comparison of the heat extraction point in the cycle, C2S, is accomplished by implementing C2S in different locations around the turbine in the C-LFR-ON non-charging cycle configuration; C-LFR-PRE has the turbine prior to C2S, C-LFR-POST has the turbine after C2S, C-LFR-PAR has the turbine in parallel to C2S, and C-LFR-CIRC uses a circulating loop instead of in-flow implementation. C-LFR-ON is the configuration used for these studies because during charging operation, flow through the CSP HX is deactivated, effectively making all non-charging cycles take the identical form of C-LFR-ON. To quantify the effectiveness of TES charging techniques, Equation \ref{eq-eta-heatstorage} defines the heat storage efficiency, $\eta_{heatstorage}$.  
\bl{explain that as a (re)compression cycle is still being run for the LFR and the components are not perfectly efficient, it will consume power moving the fluid around that needs to be balanced with some of the thermal output -BW}

@@

The high temperature sCO$_2$ leaving the LFR HX flows through the LFR turbine converting thermal energy to usable work. The outlet temperature of the LFR turbine is at a temperature suitable to charge the hot TES, therefore the flow is passed through C2S exchanging heat to the cold solar salt. The diagram outlining this process is C-LFR-PRE in Fig. \ref{c-lfr-pre}. 
\mw{Careful here. Be precise about what ``excess'' means in this context. Any fluid above absolute zero technically has excess thermal energy if the reference is zero. Depending on this definition, not all excess energy from the turbine is stored in TES. -BW} 
\mw{explain in a bit more detail what is happening that makes this configuration different from prior iterations. -BW}

@@

A sensitivity study \mw{this is one study with two parameter values, right? -BW} with varying cold CSP TES temperature is carried out to determine the impact on heat storage efficiency. The study considers two temperature values of $390^{\circ}$C and $440^{\circ}$C.

@@

\bl{explain why you are considering this case (avoid losing useful energy in compressor and turbine inefficiencies)} \mw{agree with Ben. Additional motivation is needed -BW}

@@

The three non-charging cycle models have a focus on producing the highest positive alternator power to heat input, or cycle efficiency. The two-cycle configuration, composed of C-LFR-ON and C-CSP-ON, has independent recompression cycles with dedicated compressors, turbine, recuperators, and precooler for the LFR and CSP. C-1HTR1T-ON, with the heat additions in parallel, and C-2HTR3T, with dedicated HTR for both heat additions, are cycles which incorporate the CSP and LFR in the same cycle. The results from studies done on these cycles are displayed in Table \ref{tab-noncharg}.

>>>>

\subsection{Cycle Component Modeling}
%======================================================================================
Components present in the cycles are modeled using various techniques and are discussed in more detail below. Turbines and compressors are analyzed using isentropic efficiencies. Counter-flow heat exchangers are modeled using the effectiveness-NTU method while simplified ``black box'' heat exchangers that use a simple energy balance for state point calculations are used in lieu of more detailed component models where data is available. The lead-cooled fast reactor is assumed to be a black box heat exchanger because of the constant heat input and state points on the sCO$_2$ inlet and outlet are provided. The molten salt loop for the CSP is modeled with necessary components including hot and cold TES, receiver, pumps and counter-flow heat exchangers.

@@

@@

@@

@@

@@

@@

The C2S heat exchanger is active in the 'charging' cycle operating modes, when the focus is on heat storage for later use. Pump 1 is actively moving solar salt from cold TES to hot TES through the C2S heat exchanger extracting heat from the sCO$_2$ Brayton cycle. Additionally, while the focus is on heat storage, and the heliostat field is inputting heat, pump 2 is actively transporting solar salt through the receiver to be stored in the hot TES.

@@

\subsection{Standardization of Cycle Modeling}
%======================================================================================

@@

In addition to standardized parameters, all cycles have identical recompression sides. The recompression side contains a precooler, low temperature recuperator, and two compressors; main compressor and recompressor. 
The modeled cycles are summarized in Table \ref{tab-cycle_sum}.

@@

The cycle that is using the LFR heat input in the two-cycle configuration is labeled as C-LFR-ON and the cycle diagram is illustrated in Figure \ref{c-lfr-on}.

@@

@@

The C2S is located around the turbine and LFR at state points 5 to 6 depending on the charging cycle configuration: pre, parallel, post or circulator. In the C-1HTR1T-ON cycle, the LFR HX and CSP HX have identical inlet temperatures due to splitting the flow prior to their parallel orientation. Therefore, three sensitivity studies are done on the model. The initial two studies have the low LFR temperature constrained to the value of $400^{\circ}$C with varied cold CSP TES temperature and maximized cycle efficiency. To achieve a maximum cycle efficiency, the split fraction amount of flow to the main compressor, $y_{1}$, is parametrically studied.
Two cold TES temperatures are tested with constrained LFR low temperature of $400^{\circ}$C: 
\begin{itemize}
    \item	$410^{\circ}$C: Lowest cold TES temperature possible due to the sCO$_2$ cold inlet constrained from the LFR to $400^{\circ}$C and the addition of $10^{\circ}$C approach temperature;
    \item	$440^{\circ}$C: Upper bound temperature on cold TES storage;
    \item	$390^{\circ}$C: Unconstrained LFR cold inlet temperature allows a lower cold TES of $390^{\circ}$C to be achieved. This allows for a larger temperature drop across the CSP HX, increasing dispatchability.
\end{itemize}

@@

This cycle, C-2HTR3T-ON, can be seen in Figure \ref{c-2htr3t-on} and has two high temperature recuperators and three turbines. The LFR is powering one turbine, T1, and transferring unused heat to the flow entering LFR HX through a dedicated high temperature recuperator, HTR. The cycle with heat addition from the CSP powers the other two turbines, which for modeling purposes are combined into a single turbine T2, while having a dedicated high temperature recuperator, HTR2. The two turbines displayed as T2 can be modeled as a singular turbine because their isentropic efficiencies are identical causing the inlet and outlet conditions of the turbines to be consistent with a singular turbine. Additionally, the power produced by the two turbines is proportional to mass flow rate, each receives a fraction of the mass flow rate therefore producing the same fraction of power. Summing these power fractions together yields the total power of a singular turbine in the same position. After the high temperature recuperators, the two flows are combined and sent to the LTR hot side.

@@

Charging cycle configurations accommodate energy storage modes of operation. These configurations examine the location of LFR heat extraction via C2S. To maximize the available heat for extraction, alternator net power is set to zero, therefore requiring the turbine power to balance with the compressors' demand. Despite the components being non-ideal and consuming power, the recompression cycle continues to operate, ensuring that there is mass flow to transfer heat from the Brayton cycle to C2S. The excess energy from the LFR is thermally stored in the TES for later use when grid demand increases. Comparison of the heat extraction point in the cycle, C2S, is accomplished by implementing C2S in different locations around the turbine in the C-LFR-ON non-charging cycle configuration; C-LFR-PRE has the turbine prior to C2S, C-LFR-POST has the turbine after C2S, C-LFR-PAR has the turbine in parallel to C2S, and C-LFR-CIRC uses a circulating loop instead of in-flow implementation. C-LFR-ON is the configuration used for these studies because during charging operation, flow through the CSP HX is deactivated, effectively making all non-charging cycles take the identical form of C-LFR-ON. To quantify the effectiveness of TES charging techniques, Equation \ref{eq-eta-heatstorage} defines the heat storage efficiency, $\eta_{heatstorage}$.

@@

The high temperature sCO$_2$ leaving the LFR HX flows through the LFR turbine converting thermal energy to usable work. The outlet temperature of the LFR turbine is at a temperature suitable to charge the hot TES, therefore the flow is passed through C2S exchanging heat to the cold solar salt. The diagram outlining this process is C-LFR-PRE in Fig. \ref{c-lfr-pre}.

@@

A sensitivity study with varying cold CSP TES temperature is carried out to determine the impact on heat storage efficiency. The study considers two temperature values of $390^{\circ}$C and $440^{\circ}$C.

@@

@@

The three non-charging cycle models have a focus on producing the highest positive alternator power to heat input, or cycle efficiency. The two-cycle configuration, composed of C-LFR-ON and C-CSP-ON, has independent recompression cycles with dedicated compressors, turbine, recuperators, and pre-cooler for the LFR and CSP. C-1HTR1T-ON, with the heat additions in parallel, and C-2HTR3T, with dedicated HTR for both heat additions, are cycles which incorporate the CSP and LFR in the same cycle. The results from studies done on these cycles are displayed in Table \ref{tab-noncharg}.
